HTW

### 2023-10-26 Manuscript/manuscript.qmd
update manuscript with intro

---
title: "HTW"
short-title: "Variability and Extrapolation"
date: 10-11-2023
author:
- name: Thomas Gorman
  affiliation: Indiana University
  url: www.tegorman13.github.io
  email: tegorman@iu.edu
  orcid: 0000-0001-5366-5442
- name: Rob Goldstone
  affiliation: Indiana University
  url: www.pc.cogs.indiana.edu 
  email: rgoldsto@indiana.edu
  orcid: 0000-0001-8357-8358
abstract: |
  In project 1, we applied model-based techniques to quantify and control for the similarity between training and testing experience, which in turn enabled us to account for the difference between varied and constant training via an extended version of a similarity based generalization model. In project 2, we will go a step further, implementing a full process model capable of both 1) producing novel responses and 2) modeling behavior in both the learning and testing stages of the experiment. Project 2 also places a greater emphasis on extrapolation performance following training - as varied training has often been purported to be particularly beneficial in such situations. 
keywords:
  - Learning Generalization
  - Function Learning
  - Visuomotor learning
  - Training Variability
code-repo: "Access the code, data, and analysis at <https://github.com/tegorman13/htw>"
bibliography: ../Assets/Bib/Dissertation.bib
link-citations: true
keep-md: true
---

@@

```{r setup, include=FALSE}
here::set_here(path='..')
source(here::here("Functions", "packages.R"))
test <- readRDS(here("data/e1_08-21-23.rds")) |> filter(expMode2 == "Test") 
#options(brms.backend="cmdstanr",mc.cores=4)
e1Sbjs <- test |> group_by(id,condit) |> summarise(n=n())
testAvg <- test %>% group_by(id, condit, vb, bandInt,bandType,tOrder) %>%
  summarise(nHits=sum(dist==0),vx=mean(vx),dist=mean(dist),sdist=mean(sdist),n=n(),Percent_Hit=nHits/n)

@@

```{r}
#| label: tbl-e1-test-nf-deviation
#| tbl-cap: "Testing Deviation - Empirical Summary"
#| tbl-subcap:
#|   - "Full datasets"
#|   - "Intersection of samples with all labels available"
#| layout-ncol: 1
#| include: true
#| echo: false

@@

result$constant |>kbl(booktabs = TRUE,
                        linesep = "\\addlinespace[0.5em]",
                        caption = paste("Summary of Deviation- Constant"))
# |>
#   kable_styling(font_size = ifelse(fmt_out == "latex", 8.5, NA))

result$varied |> kbl(booktabs = TRUE,
                        linesep = "\\addlinespace[0.5em]",
                        caption = paste("Summary of Deviation- Varied"))

@@

```{r}
#| label: tbl-e1-bmm-dist
#| tbl-cap: "Experiment 1. Bayesian Mixed Model predicting absolute deviation as a function of condition (Constant vs. Varied) and Velocity Band"
#| tbl-subcap: ["Constant Testing1 - Deviation", "Varied Testing - Deviation"]
#| include: true
#| echo: false
#| eval: true
modelName <- "e1_testDistBand_RF_5K"
e1_distBMM <- brm(dist ~ condit * bandInt + (1 + bandInt|id),
                      data=test,file=paste0(here::here("data/model_cache",modelName)),
                      iter=5000,chains=4)
GetModelStats(e1_distBMM) |> kbl(booktabs = TRUE,caption = paste("Coefficients"))

e1_distBMM |> 
  emmeans("condit",by="bandInt",at=list(bandInt=c(100,350,600,800,1000,1200)),
          epred = TRUE, re_formula = NA) |> 
  pairs() |> gather_emmeans_draws()  |> 
   summarize(median_qi(.value),pd=sum(.value>0)/n()) |>
   select(contrast,Band=bandInt,value=y,lower=ymin,upper=ymax,pd) |> 
   mutate(across(where(is.numeric), \(x) round(x, 2)),
          pd=ifelse(value<0,1-pd,pd)) |> kbl(booktabs = TRUE)
# |> 
#   kbl(caption="Contrasts")

@@

```{r}
#| label: tbl-e1-test-nf-vx
#| tbl-cap: "Testing vx - Empirical Summary"
#| layout-ncol: 1
#| include: true

result <- test_summary_table(test, "vx","X Velocity", mfun = list(mean = mean, median = median, sd = sd))
result$constant |> kbl(booktabs = TRUE)
result$varied |> kbl(booktabs = TRUE)

@@

```{r}
#| label: tbl-e1-bmm-vx
#| tbl-cap: "Experiment 1. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band"
#| tbl-subcap: ["Model fit to all 6 bands", "Model fit to 3 extrapolation bands"]
#| include: true
#| eval: true
e1_vxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                        data=test,file=paste0(here::here("data/model_cache", "e1_testVxBand_RF_5k")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))
GetModelStats(e1_vxBMM ) |> kbl(booktabs=T, caption="Fit to all 6 bands")

@@

  indvSlopes |> mutate(Condition=condit) |>  group_by(Condition) |> 
    reframe(enframe(quantile(SlopeInt, c(0.0,0.25, 0.5, 0.75,1)), "quantile", "SlopeInt")) |> 
  pivot_wider(names_from=quantile,values_from=SlopeInt,names_prefix="Q_") |>
  group_by(Condition) |>
  summarise(across(starts_with("Q"), list(mean = mean))) |> kbl()

@@

```{r}
#| label: fig-e1-indv-slopes
#| fig-cap: Subset of Varied and Constant Participants with the smallest and largest estimated slope values. Red lines represent the best fitting line for each participant, gray lines are 200 random samples from the posterior distribution. Colored points and intervals at each band represent the empirical median and 95% HDI. 
#| fig-subcap: ["subset with largest slopes", "subset with smallest slopes"]
#| fig-height: 10
#| fig-width: 12
#| include: true

@@

result <- test_summary_table(testE2, "dist","Deviation", mfun = list(mean = mean, median = median, sd = sd))
result$constant |> kbl()
result$varied |> kbl()
# make kable table with smaller font size
# result$constant |> kbl(caption="Constant Testing - Deviation",booktabs=T,escape=F) |> kable_styling(font_size = 7)
```

@@

```{r}
#| label: tbl-e2-bmm-dist
#| tbl-cap: "Experiment 2. Bayesian Mixed Model predicting absolute deviation as a function of condition (Constant vs. Varied) and Velocity Band"
#| eval: true

@@

e2_distBMM |> 
  emmeans("condit",by="bandInt",at=list(bandInt=c(100,350,600,800,1000,1200)),
          epred = TRUE, re_formula = NA) |> 
  pairs() |> gather_emmeans_draws()  |> 
   summarize(median_qi(.value),pd=sum(.value>0)/n()) |>
   select(contrast,Band=bandInt,value=y,lower=ymin,upper=ymax,pd) |> 
   mutate(across(where(is.numeric), \(x) round(x, 2)),
          pd=ifelse(value<0,1-pd,pd)) |>
   kbl(caption="Contrasts")

@@

@@

resultOrig <- test_summary_table(testE3 |> filter(bandOrder=="Original"), "dist","Deviation", mfun = list(mean = mean, median = median, sd = sd))
resultOrig$constant |> kbl()
resultOrig$varied |> kbl()

resultRev <- test_summary_table(testE3 |> filter(bandOrder=="Reverse"), "dist","Deviation", mfun = list(mean = mean, median = median, sd = sd))
resultRev$constant |> kbl()
resultRev$varied |> kbl()

@@

```{r}
#| label: fig-e3-test-vx
#| fig-cap: e3 testing x velocities. Translucent bands with dash lines indicate the correct range for each velocity band. 
#| fig-width: 14
#| fig-height: 15
#| column: screen-inset-right
# testE3 |> filter(bandOrder=="Original")|> group_by(id,vb,condit) |> plot_distByCondit()
# testE3 |> filter(bandOrder=="Reverse")|> group_by(id,vb,condit) |> plot_distByCondit() +ggtitle("test")

@@

@@

resultOrig <- test_summary_table(testE3 |> filter(bandOrder=="Original"), "vx","X Velocity", mfun = list(mean = mean, median = median, sd = sd))
resultOrig$constant |> kbl()
resultOrig$varied |> kbl()

resultRev <- test_summary_table(testE3 |> filter(bandOrder=="Reverse"), "vx","X Velocity", mfun = list(mean = mean, median = median, sd = sd))
resultRev$constant |> kbl()
resultRev$varied |> kbl()
```

>>>>

---
title: "HTW"
short-title: "Variability and Extrapolation"
date: 10-11-2023
author:
- name: Thomas Gorman
  affiliation: Indiana University
  url: www.tegorman13.github.io
  email: tegorman@iu.edu
  orcid: 0000-0001-5366-5442
- name: Rob Goldstone
  affiliation: Indiana University
  url: www.pc.cogs.indiana.edu 
  email: rgoldsto@indiana.edu
  orcid: 0000-0001-8357-8358
abstract: |
  In project 1, we applied model-based techniques to quantify and control for the similarity between training and testing experience, which in turn enabled us to account for the difference between varied and constant training via an extended version of a similarity based generalization model. In project 2, we will go a step further, implementing a full process model capable of both 1) producing novel responses and 2) modeling behavior in both the learning and testing stages of the experiment. Project 2 also places a greater emphasis on extrapolation performance following training - as varied training has often been purported to be particularly beneficial in such situations. 
keywords:
  - Learning Generalization
  - Function Learning
  - Visuomotor learning
  - Training Variability
code-repo: "Access the code, data, and analysis at <https://github.com/tegorman13/htw>"
bibliography: ../Assets/Bib/Dissertation.bib
link-citations: true
---

@@

```{r setup, include=FALSE}
here::set_here(path='..')

source(here::here("Functions", "packages.R"))

# pacman::p_load(tidyverse,knitr,tidybayes,brms, lme4, bayesplot,bayestestR,parameters,marginaleffects,
#                 emmeans, equatiomatic, here, pacman,  broom,
#                broom.mixed,lme4,emmeans,here,
#                 wesanderson,glue, ggdist,ggforce,patchwork,gghalves,ggh4x,
#                 install = TRUE,
#                 update = FALSE
#                )
# walk(c(here("Functions/Display_Functions.R"), here("Functions/org_functions.R"), 
#        here("Functions/Table_Functions.R")), source)

test <- readRDS(here("data/e1_08-21-23.rds")) |> filter(expMode2 == "Test") 
#options(brms.backend="cmdstanr",mc.cores=4)
e1Sbjs <- test |> group_by(id,condit) |> summarise(n=n())
testAvg <- test %>% group_by(id, condit, vb, bandInt,bandType,tOrder) %>%
  summarise(nHits=sum(dist==0),vx=mean(vx),dist=mean(dist),sdist=mean(sdist),n=n(),Percent_Hit=nHits/n)

@@

```{r}
#| label: tbl-e1-test-nf-deviation
#| tbl-cap: "Testing Deviation - Empirical Summary"
#| tbl-subcap:
#|   - "Full datasets"
#|   - "Intersection of samples with all labels available"

@@

result$constant |>kable(booktabs = TRUE,
                        linesep = "\\addlinespace[0.5em]")
                        #caption = paste("Summary of Deviation- Constant"))
# |>
#   kable_styling(font_size = ifelse(fmt_out == "latex", 8.5, NA))

result$varied |> kable(booktabs = TRUE,
                        linesep = "\\addlinespace[0.5em]")
                        #caption = paste("Summary of Deviation- Varied"))

@@

```{r}
#| label: tbl-e1-bmm-dist
#| tbl-cap: "Experiment 1. Bayesian Mixed Model predicting absolute deviation as a function of condition (Constant vs. Varied) and Velocity Band"
#| tbl-subcap: ["Constant Testing1 - Deviation", "Varied Testing - Deviation"]

modelName <- "e1_testDistBand_RF_5K"
e1_distBMM <- brm(dist ~ condit * bandInt + (1 + bandInt|id),
                      data=test,file=paste0(here::here("data/model_cache",modelName)),
                      iter=5000,chains=4)
GetModelStats(e1_distBMM) |> kable(booktabs = TRUE,caption = paste("Coefficients"))

e1_distBMM |> 
  emmeans("condit",by="bandInt",at=list(bandInt=c(100,350,600,800,1000,1200)),
          epred = TRUE, re_formula = NA) |> 
  pairs() |> gather_emmeans_draws()  |> 
   summarize(median_qi(.value),pd=sum(.value>0)/n()) |>
   select(contrast,Band=bandInt,value=y,lower=ymin,upper=ymax,pd) |> 
   mutate(across(where(is.numeric), \(x) round(x, 2)),
          pd=ifelse(value<0,1-pd,pd)) |> kable(booktabs = TRUE)
# |> 
#   kbl(caption="Contrasts")

@@

```{r}
#| label: tbl-e1-test-nf-vx
#| tbl-cap: "Testing vx - Empirical Summary"
#| tbl-subcap: ["Constant", "Varied"]

result <- test_summary_table(test, "vx","X Velocity", mfun = list(mean = mean, median = median, sd = sd))
result$constant |> kable(booktabs = TRUE)
result$varied |> kable(booktabs = TRUE)

@@

```{r}
#| label: tbl-e1-bmm-vx
#| tbl-cap: "Experiment 1. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band"
#| tbl-subcap: ["Model fit to all 6 bands", "Model fit to 3 extrapolation bands"]
#| include: true
#| eval: true
e1_vxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                        data=test,file=paste0(here::here("data/model_cache", "e1_testVxBand_RF_5k")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))
GetModelStats(e1_vxBMM ) |> kable(booktabs=T, caption="Fit to all 6 bands")

@@

  indvSlopes |> mutate(Condition=condit) |>  group_by(Condition) |> 
    reframe(enframe(quantile(SlopeInt, c(0.0,0.25, 0.5, 0.75,1)), "quantile", "SlopeInt")) |> 
  pivot_wider(names_from=quantile,values_from=SlopeInt,names_prefix="Q_") |>
  group_by(Condition) |>
  summarise(across(starts_with("Q"), list(mean = mean))) |> kable()

@@

test

```{r}
#| label: fig-e1-indv-slopes
#| fig-cap: Subset of Varied and Constant Participants with the smallest and largest estimated slope values. Red lines represent the best fitting line for each participant, gray lines are 200 random samples from the posterior distribution. Colored points and intervals at each band represent the empirical median and 95% HDI. 
#| fig-subcap: ["subset with largest slopes", "subset with smallest slopes"]
#| fig-height: 6
#| fig-width: 9

@@

result <- test_summary_table(testE2, "dist","Deviation", mfun = list(mean = mean, median = median, sd = sd))
result$constant |> kable()
result$varied |> kable()
# make kable table with smaller font size
# result$constant |> kbl(caption="Constant Testing - Deviation",booktabs=T,escape=F) |> kable_styling(font_size = 7)
```

@@

```{r}
#| label: tbl-e2-bmm-dist
#| tbl-cap: "Experiment 2. Bayesian Mixed Model predicting absolute deviation as a function of condition (Constant vs. Varied) and Velocity Band"
#| tbl-subcap: ["Model fits", "Contrasts"]

@@

e2_distBMM |> 
  emmeans("condit",by="bandInt",at=list(bandInt=c(100,350,600,800,1000,1200)),
          epred = TRUE, re_formula = NA) |> 
  pairs() |> gather_emmeans_draws()  |> 
   summarize(median_qi(.value),pd=sum(.value>0)/n()) |>
   select(contrast,Band=bandInt,value=y,lower=ymin,upper=ymax,pd) |> 
   mutate(across(where(is.numeric), \(x) round(x, 2)),
          pd=ifelse(value<0,1-pd,pd)) |>
   kable()

@@

In addition to accuracy/deviation. We also assessed the ability of participants to reliably discriminate between the velocity bands (i.e. responding differently when prompted for band 600-800 than when prompted for band 150-350). @tbl-e2-test-nf-vx shows descriptive statistics of this measure, and Figure 1 visualizes the full distributions of throws for each combination of condition and velocity band. To quantify discrimination, we again fit Bayesian Mixed Models as above, but this time the dependent variable was the raw x velocity generated by participants.

```{r}
#| label: fig-e2-test-vx
#| fig-cap: E2 testing x velocities. Translucent bands with dash lines indicate the correct range for each velocity band. 
#| fig-width: 11
#| fig-height: 9
testE2 %>% group_by(id,vb,condit) |> plot_distByCondit()

```{r}
#| label: tbl-e2-test-nf-vx
#| tbl-cap: "Testing vx - Empirical Summary"
#| tbl-subcap: ["Constant Testing - vx", "Varied Testing - vx"]
#| layout-ncol: 2

result <- test_summary_table(testE2, "vx","X Velocity" ,mfun = list(mean = mean, median = median, sd = sd))
result$constant |> kable()
result$varied |> kable()

```{r}
#| label: tbl-e2-bmm-vx
#| tbl-cap: "Experiment 2. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band"
e2_vxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                        data=testE2,file=paste0(here::here("data/model_cache", "e2_testVxBand_RF_5k")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))
mt3 <-GetModelStats(e2_vxBMM ) |> kable(escape=F,booktabs=T)
mt3

cd1 <- get_coef_details(e2_vxBMM, "conditVaried")
sc1 <- get_coef_details(e2_vxBMM, "bandInt")
intCoef1 <- get_coef_details(e2_vxBMM, "conditVaried:bandInt")
```

See @tbl-e2-bmm-vx for the full model results.

When examining discrimination ability using the model predicting raw x-velocity, the results were less clear than those of the absolute deviation analysis. The slope on Velocity Band (β = `r sc1$estimate`, 95% CrI \[`r sc1$conf.low`, `r sc1$conf.high`\]) indicates that participants showed good discrimination between bands overall. However, the interaction term suggested this effect was not modulated by training condition (β = `r intCoef1$estimate`, 95% CrI \[`r intCoef1$conf.low`, `r intCoef1$conf.high`\]) Thus, while varied training may provide some advantage for accuracy, both training conditions seem to have similar abilities to discriminate between velocity bands.

```{r}
#| label: fig-e2-bmm-vx
#| fig-cap: Conditional effect of training condition and Band. Ribbons indicate 95% HDI.

e2_vxBMM |> emmeans( ~condit + bandInt, 
                       at = list(bandInt = c(100, 350, 600, 800, 1000, 1200))) |>
  gather_emmeans_draws() |>
  ggplot(aes(x = bandInt, y = .value, color = condit, fill = condit)) + 
  stat_dist_pointinterval() +
  stat_lineribbon(alpha = .25, size = 1, .width = c(.95)) +
  ylab("Predicted X Velocity") + xlab("Band")+
  scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                     labels = levels(testE2$vb), 
                     limits = c(0, 1400)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) 
```

@@

resultOrig <- test_summary_table(testE3 |> filter(bandOrder=="Original"), "dist","Deviation", mfun = list(mean = mean, median = median, sd = sd))
resultOrig$constant |> kable()
resultOrig$varied |> kable()

resultRev <- test_summary_table(testE3 |> filter(bandOrder=="Reverse"), "dist","Deviation", mfun = list(mean = mean, median = median, sd = sd))
resultRev$constant |> kable()
resultRev$varied |> kable()

@@

```{r}
#| label: fig-e3-test-vx
#| fig-cap: e3 testing x velocities. Translucent bands with dash lines indicate the correct range for each velocity band. 
#| fig-width: 12
#| fig-height: 13

# testE3 |> filter(bandOrder=="Original")|> group_by(id,vb,condit) |> plot_distByCondit()
# testE3 |> filter(bandOrder=="Reverse")|> group_by(id,vb,condit) |> plot_distByCondit() +ggtitle("test")

@@

# column: screen-inset-right

@@

resultOrig <- test_summary_table(testE3 |> filter(bandOrder=="Original"), "vx","X Velocity", mfun = list(mean = mean, median = median, sd = sd))
resultOrig$constant |> kable()
resultOrig$varied |> kable()

resultRev <- test_summary_table(testE3 |> filter(bandOrder=="Reverse"), "vx","X Velocity", mfun = list(mean = mean, median = median, sd = sd))
resultRev$constant |> kable()
resultRev$varied |> kable()
```
